---
title: Linear Regression in R 
author: Alessia De Biase and Bruno Barakat
package: lab4ab
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `lab4ab` package provides the main function to compute a linear regression when a set of data and the variables to investigate are given. 
Compared to the classic linear regression function (`lm()`) `linreg()` helps to calculate all the most significative elements of a linear regression in just one step just giving the observations (contained in a dataset) and the formula. `lab4ab` also provides a plotting function and a summary function just as `lm()` does.

## Linear regression function: `linreg()`

`linreg(data, formula)` needs two arguments:

- *formula* : the linear regression formula the user wants to compute, where on the left is the dependent variable and on the right all the indipendent variables (for example: `Petal.Length ~ Sepal.Width + Sepal.Length`)
- *data* : a dataframe containing all the features (one for each column);

Applying the function to those objects the dependent variable is stored in a vector *Y* and the independent variables in a matrix *X*. At the same time all the following statistics are calculated:

1. **Regression coefficients** (`beta_hat`) : $\hat{\beta}= (X^TX)^{-1}X^Ty$

2. **The fitted values** (`fitted`) : $\hat{y}= X\hat{\beta}$

3. **The residuals** (`res`) : $\hat{e}= y-X\hat{\beta}$

4. **The degree of freedom** (`df`) : $df=n-p$ , where $n$ is the number of observations and $p$ is the number of parameters in the model

5. **The residual variance** (`res_var`) : $\hat{\sigma}^2=\frac{e^Te}{df}$

6. **The variance of the regression coefficients** (`reg_var`) : $\hat{Var}(\hat{\beta})=\hat{\sigma}^2(X^TX)^{-1}) $

7. **The t-values for each coefficient** (`t_value`) :$t_\beta=\frac{\hat{\beta}}{\sqrt{Var(\hat{\beta})}}$

8. **The p-values for each coefficient** (`p_value`) : $P(|>t|)$ the probability of each coefficient to be equal to 0

*linreg* is a class, all the calculations have been stored in it, so to access to one of the previews statistics it may be used *$* as it follows:

```{r}
data <- iris
formula <- Petal.Length ~ Sepal.Width + Sepal.Length

my_linear_reg <- linreg(formula, data)

residuals <- my_linear_reg$res

residuals 

```


The package provides also three other methods which return some of the statistics showed before: 

- `$resid()`: which returns the vector of residuals (point 3.)
- `$pred()` : which returns the predicted values (point 2.)
- `$coef()` : which returns the regression coefficients with their names (point 1.)


## The printing method:  `print()`

The `$print()` shows exactly the same data of `lm()` function: it provides the coefficients found in the linear regression and their names. 

```{r}
data <- iris

formula <- Petal.Length ~ Sepal.Width + Sepal.Length

my_linear_reg <- linreg(formula, data)

my_linear_reg$print()

```

## Plotting

In `lab4ab` package there is also a method which contains two graphs: one representing the residuals vs the fitted values, the second one a scaled version of the first one, the residuals have been standardized by the formula: $$\sqrt{ \frac{|mean(res)-res|}{\sqrt{{Var(res)}}}}$$ .
The line represents the trend of the observations in the graphs. The smooth has been realized with a local polynomial regression fitting with a span (parameter which controls the degree of smoothing) of 1.5. It looks very close to the one obtained by `plot(lm())`. The methid `$plot()` plots, using *ggplot2*  package that has to be installed before using `lab4ab`. The graphs respect some of the graphical criteria of Linkoping University.

```{r message=FALSE, fig.align="center",fig.height = 5, fig.width = 9}
my_linear_reg$plot()

```

## `$summary()`

Last method in the package is `summary()`. It's aim is to return a similar printout as printed for `lm()`. The statistics of this output are: 

```{r message=FALSE}

my_linear_reg$summary()

```

## Examples

```{r}
data <- iris

formula <- Petal.Length ~ Species

my_linear_reg <- linreg(formula, data)

```

```{r, echo=FALSE, results='asis'}
knitr::kable(head(iris,10))
```
```{r}

my_linear_reg$print()

```
```{r}

my_linear_reg$summary()

```





